Chromosomal duplication can be an evolutionary detour on the path to adaptation

### 2022-09-13 ms/ms.tex
revise following IK comments:

However, a model without aneuploidy (where the aneuploidy rate is fixed at zero, $\delta=0$), fails to explain the experimental observations~(Figure~\ref{fig:2n*-fixation}). %% generated with MAP-and-hdi.ipynb
The estimated mutation rate without aneuploidy is $\mu=7.98\cdot10^{-9}\ [7.906\cdot10^{-9}-8.138\cdot10^{-9}]$, much lower compared to a model with aneuploidy and suggesting a target size of just 40. 
The fitness of the mutant is also much lower at $w_{\eumt}=1.013\ [1.012-1.013]$.
This is because, without aneuploidy, a high mutation rate or fitness effect will lead to faster appearance and fixation of \eumt\ than in the experimental observations. Even with these lower estimates, the model fit is worse than that of a model with aneuploidy~(\Cref{fig:2n*-fixation}).

@@

\paragraph{Sensitivity analysis.} 
Changing a single parameter while keeping the rest fixed at the MAP estimate produces a worse fit to the data~(\Cref{fig:sensitivity}).
Furthermore, we fitted models with a mutation rate fixed at $\mu$=$10^{-5}$, $10^{-6}$ and $10^{-7}$.
We inferred similar parameters estimates for the model with $\mu=10^{-6}$ compared to the model with a free $\mu$ parameter, in which the inferred mutation rate is $\mu \approx 3\cdot10^{-6}$.
Models with $\mu$=$10^{-5}$ and $\mu$=$10^{-7}$ inferred different parameters estimates, including higher aneuploidy rate and the model with $\mu$=$10^{-5}$ has lower fitness value for \anwt\ (\Cref{fig:mu}).

@@

% generated with alt-prior.ipynb
We estimated the parameters under this extended informative prior.
Inference took much longer to run but the posterior distribution seemed to converge, as it did not change much in the final iterations. 
The posterior predictive plot shows that inference with this extended prior produces a posterior distribution that fails to explain the empirical observations~(pink in Figure~\ref{fig:2n*-fixation}).
However, the inferred posterior distribution is considerably narrower (compare \Cref{fig:posterior,fig:posterior-alt}) and therefore parameter estimates are less variable.
The estimated mutation rate was much lower compared to the main informative prior, 
% generated with MAP-and-hdi.ipynb
with $\mu=2.474\cdot10^{-9}\ [2.423\cdot10^{-9}-2.612\cdot10^{-9}]$. Other parameter estimates are: $$\delta=2.705\cdot10^{-3}\ [2.094\cdot10^{-3}-3.094\cdot10^{-3}]$,
$w_{\anwt}=1.022\ [1.021-1.024]$,
$w_{2n+1^*}=1.052\ [1.05-1.054]$,
$w_{2n^*}=1.053\ [1.051-1.055]$. 
Notably, the maximum posterior ratio $w_{\eumt}/w_{\anwt}=1.0009$ is much lower than the maximum prior ratio of $1.033$ (\Cref{fig:growth-curves}H) and closer to the ratio of $1$ that we assume in our standard prior.
Together with the posterior predictive results, we conclude that the main informative prior is preferable over the extended informative prior.

\subsection*{Supplementary Results}

>>>>

However, a model without aneuploidy (where the aneuploidy rate is fixed at zero, $\delta=0$), fails to explain the experimental observations~(Figure~\ref{fig:2n*-fixation}). %% generated with MAP-and-hdi.ipynb
The estimated mutation rate without aneuploidy is $\mu=7.98\cdot10^{-9}\ [7.906\cdot10^{-9}-8.138\cdot10^{-9}]$, much lower compared to a model with aneuploidy and suggesting a target size of just 40. 
The fitness of the mutant is also much lower at $w_{\eumt}=1.013\ [1.012-1.013]$.
This is because, without aneuploidy, a high mutation rate or fitness effect will lead to faster appearance and fixation of \eumt\ than in the experimental observations.

@@

\paragraph{Sensitivity analysis.} 
Changing a single parameter while keeping the rest fixed at the MAP estimate produces a worse fit to the data~(\Cref{fig:sensitivity}).
Furthermore, we fitted models with a mutation rate fixed at $\mu=10^{-5}$, $10^{-6}$ and $10^{-7}$.
We inferred similar parameters estimates for the model with $\mu=10^{-6}$ compared to the model with a free $\mu$ parameter, in which the inferred mutation rate is $\mu \approx 3\cdot10^{-6}$.
Inference assuming $\mu=10^{-5}$ or $\mu=10^{-7}$ produced a higher estimate of the aneuploidy rate, $\delta$; assuming $\mu=10^{-7}$, the estimated fitness of \anwt\ was lower; otherwise, estimates were similar to our main model (\Cref{fig:mu}).

@@

% generated with alt-prior.ipynb
We estimated the parameters under this extended informative prior.
Inference took much longer to run but the posterior distribution seemed to converge, as it did not change much in the final iterations. 
The posterior predictive plot shows that inference with this extended prior produces a posterior distribution that fails to explain the empirical observations~(pink in Figure~\ref{fig:2n*-fixation}).
However, the inferred posterior distribution is considerably narrower (compare \Cref{fig:posterior,fig:posterior-alt}) and therefore parameter estimates are less variable.
The estimated mutation rate was much lower compared to the main informative prior, 
% generated with MAP-and-hdi.ipynb
with $\mu=2.474\cdot10^{-9}\ [2.423\cdot10^{-9}-2.612\cdot10^{-9}]$. Other parameter estimates are: $\delta=2.705\cdot10^{-3}\ [2.094\cdot10^{-3}-3.094\cdot10^{-3}]$,
$w_{\anwt}=1.022\ [1.021-1.024]$,
$w_{2n+1^*}=1.052\ [1.05-1.054]$,
$w_{2n^*}=1.053\ [1.051-1.055]$, the latter two being much higher compare to the main informative prior. 
Notably, the mode of the posterior ratio $w_{\eumt}/w_{\anwt}=1.0009$ is much lower than the mode of the prior ratio of $1.033$ (\Cref{fig:growth-curves}H) and closer to the ratio of $1$ that we assume in the main informative prior.
Together with the posterior predictive results, we conclude that the main informative prior is preferable over the extended informative prior.

\paragraph{Model with transitions to less-fit genotypes}
